---
title: "PS-1 Markdown"
author: "Nicole Bohra"
date: "1/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("VGAM")
```

#Section 1: Outliers in Different Distributions
###1.1: Normal Distribution
1. Setting up R and N
``` {r one, include = TRUE}
r <- 1000
n <- 1000
```
2. R means of N standard norms
``` {r sapply, include = TRUE}
x <- sapply(1:n, function(s) mean(rnorm(n)))
```
3. 
Mean of means: 
```{r means, include = TRUE} 
mean(mean(x))
```
Standard Deviation: 
```{r sd, include = TRUE}
sd(x)
```
Confidence Interval: 
```{r interval, include=TRUE}
left <- quantile(x, 0.025)
left
right <- quantile(x, 0.975)
right
```

4. Based on the theoretical considerations, with an increase in sample sizes we can expect the confidence interval to decrease since it decreases the standard error. The larger the sample size, the less uncertainty there is about the true population mean, and the estimate becomes more precise. In this particular scenario, I am 95% confident that the true mean of this dataset lies between 6.190 and 24594.86 with a sample size of 1,000.

5. Different Sample Sizes
```{r other, include=TRUE}
r2 <- 1000
n2 <- 10
x2 <- sapply(1:n2, function(s) mean(rnorm(n2)))
mean(mean(x2))
sd(x2)
left <- quantile(x2, 0.025)
left
right <- quantile(x2, 0.975)
right

```
With a sample size decreased to 10, we see that the confidence interval (still at 95%) has a wider range, with values from -0.304 to 0.496.

###1.2: Pareto Distribution
1. K Values in Pareto
```{r pareto, echo=FALSE}
n <- 10000
k1 <- 100
k2 <- 10
k3 <- 1

s1 <- VGAM::rpareto(n, 1, k1)
m1 <- mean(s1)
{hist(s1)
abline(v = m1, col = "blue", lwd = 2)}

s2 <- VGAM::rpareto(n, 1, k2)
m2 <- mean(s2)
{hist(s2)
abline(v = m2, col = "red", lwd = 2)}

s3 <- rpareto(n, 1, k3)
m3 <- mean(s3)
{hist(s3)
abline(v = m3, col = "green", lwd = 2)}
```
2. There is more of a distribution shown when k = 10 and 100 versus when k = 1. The distribution is more spread for when k = 100 than the other k values, however both k = 10 and 100 histograms show a large right skew and a heavy tail. The shapes of these histograms do not follow a normal distribution. A high proportion of the graph has low values, whereas a low proportion of the graph has high values.

3. Different K and N values
```{r pareto2, include=TRUE}
k_less <- 0.5
k_greater <- 2
r <- 1000
n1 <- 10
n2 <- 1000
n3 <- 10000

# Sample size is 10, and k is 0.5.
x_pareto_less <- sapply(1:n1, function(s) mean(rpareto(n1, 1, k_less)))
mean_less <- mean(mean(x_pareto_less))
sd(x_pareto_less)
left_less <- quantile(x_pareto_less, 0.025)
right_less <- quantile(x_pareto_less, 0.975)
left_less
right_less

# Sample size is 10, and k is 2.
x_pareto_greater <- sapply(1:n1, function(s) mean(rpareto(n1, 1, k_greater)))
mean_greater <- mean(mean(x_pareto_greater))
sd(x_pareto_greater)
left_greater <- quantile(x_pareto_greater, 0.025)
right_greater <- quantile(x_pareto_greater, 0.975)
left_greater
right_greater

# Sample size is 1000, and k is 0.5.
x_pareto_less_2 <- sapply(1:n2, function(s) mean(rpareto(n2, 1, k_less)))
mean_less_2 <- mean(mean(x_pareto_less_2))
sd(x_pareto_less_2)
left_less_2 <- quantile(x_pareto_less_2, 0.025)
right_less_2 <- quantile(x_pareto_less_2, 0.975)
left_less_2
right_less_2

# Sample size is 1000, and k is 2. 
x_pareto_greater_2 <- sapply(1:n2, function(s) mean(rpareto(n2, 1, k_greater)))
mean_greater_2 <- mean(mean(x_pareto_greater_2))
sd(x_pareto_greater_2)
left_greater_2 <- quantile(x_pareto_greater_2, 0.025)
right_greater_2 <- quantile(x_pareto_greater_2, 0.975)
left_greater_2
right_greater_2
```
4. From changing the sample sizes and lower bounds in the Pareto distribution, my findings show that the standard deviations and confidence regions (95%) vary widely. In the case of a normal distribution, the sample size affects the confidence interval; the smaller the sample size, the wider the confidence interval gets. With the Pareto distribution, the sample size does not seem to affect it as much as the k-value. When sampling with n = 10 and 1000, the standard deviations and confidence regions stayed relatively the same. However, when k was increased to 2, the standard deviation was much smaller and the confidence interval was narrow. (I was unable to try for k = 10000 due to runtime errors.)

#Section 2: Linear Transformation of Images
###2.1: Rotate Matrices
1. Rotate Matrix A
```{r matA, include=TRUE}
A <- matrix(c(0,0, 0,2, 1,1), 3, 2, byrow = TRUE)
plot(A[,1], A[,2], type ="l", asp=1)
Rot <- function(alpha) {
  a <- alpha*pi/180
  matrix(c(cos(a), sin(a), -sin(a), cos(a)), 2, 2)
}

A_1 <- A%*%Rot(45)
plot(A_1[,1], A_1[,2], type ="l", asp=1)

A_2 <- A%*%Rot(-225)
plot(A_2[,1], A_2[,2], type ="l", asp=1)

A_3 <- A%*%Rot(60)
plot(A_3[,1], A_3[,2], type ="l", asp=1)

```

2. Create new matrix object "B"
```{r matB, include=TRUE}
B <- matrix(c(2,3, 5,6, 10,-12), 3, 2, byrow = TRUE)
B
```

3. Rotate Matrix B
```{r rotate, include=TRUE}
plot(B[,1], B[,2], type="b", asp=1 )

B_1 <- B%*%Rot(60)
plot(B_1[,1], B_1[,2], type="b", asp=1)

B_2 <- B%*%Rot(120)
plot(B_2[,1], B_2[,2], type="b", asp=1)

B_3 <- B%*%Rot(-180)
plot(B_3[,1], B_3[,2], type="b", asp=1)
```

###2.2: Rotate Data
1. Read in Crazy Hat file
```{r read, include=TRUE}
setwd("/Users/nicolebohra/INFO371")
data <- read.table('crazy_hat.tsv', header = TRUE)
```

2. Plot graph and connect lines
```{r plot, include=TRUE}
ggplot(data, aes(x = x, y = y, group = group)) + geom_path()
```

3. Rotate Crazy Hat matrix object
```{r transform, include=TRUE}
mat <- data.matrix(data)
Rot <- function(alpha) {
  a <- alpha*pi/180
  matrix(c(cos(a), sin(a), -sin(a), cos(a)), 2, 2)
}
transMatrix1 <- mat[,1:2]%*%Rot(60)
transMatrix1.df <- cbind(transMatrix1[,1:2], mat[,3])
transMatrix1.df <- as.data.frame(transMatrix1.df)
ggplot(transMatrix1.df, aes(x = V1, y = V2, group = V3)) + geom_path()

transMatrix2 <- mat[,1:2]%*%Rot(-120)
transMatrix2.df <- cbind(transMatrix2[,1:2], mat[,3])
transMatrix2.df <- as.data.frame(transMatrix2.df)
ggplot(transMatrix2.df, aes(x=V1, y = V2, group = V3)) + geom_path()

transMatrix3 <- mat[,1:2]%*%Rot(85)
transMatrix3.df <- cbind(transMatrix3[,1:2], mat[,3])
transMatrix3.df <- as.data.frame(transMatrix3.df)
ggplot(transMatrix3.df, aes(x=V1, y = V2, group = V3)) + geom_path()
```

###2.3: Flip and Stretch
1. Flip matrix over y-axis
```{r flip, include=TRUE}
X <- matrix(c(1,0, 1,3, 2,3), 3, 2, byrow = TRUE)
plot(x = X[,1], y =X[,2], type="l", asp=1)

X_flip<- X%*%matrix(c(-1, 0, 0, 1), 2, 2)
plot(X_flip[,1], X_flip[,2], type="l", asp=1)
```

2. Stretch matrix object
```{r stretch, include=TRUE}
X_stretch <- X%*%matrix(c(1, 0, 0, 2), 2, 2) 
plot(X_stretch[,1], X_stretch[,2], type="l", asp=1)
```

3. Transform Crazy Hat object
```{r transformation, include=TRUE}
flipStretchMatrix <- function(mat, alpha, stretchFactor){
  mat <- mat[,1:2] %*% Rot(alpha)
  mat[,2] <- mat[,2] * stretchFactor
  return(mat)
}
flipStretchData <- function(data, alpha, stretchFactor){
  matrix_flip <- data.matrix(data)
  matrix_flip <- flipStretchMatrix(matrix_flip, alpha, stretchFactor)
  data$x <- matrix_flip[,1]
  data$y <- matrix_flip[,2]
  ggplot(data, aes(x, y, group=group)) + geom_path()
}
flipStretchData(data, 45, 3)
```

4. 
To modify the matrix in which post-multiply and pre-multiply outputs the same result, a conversion using the transpose function would yield the same result. The equivalent to (AB) would be to multiply the transposed B by transposed A and transpose the entire result. The equation would look like: A x B = (B' x A')'.
