---
title: "Lab 4"
author: "Nicole Bohra"
date: "2/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("dplyr")
library("dplyr")
```

##1.1 Before and After Estimator
```{r table}
data <- read.table(file = 'yrbs.tsv', header = TRUE)
data$before <- data$year < 2006
mean_before <- mean(data$before == TRUE)
mean_after <- mean(data$before == FALSE)
mean_table <- cbind(mean_before, mean_after)

regression1 <- lm(meth~before, data = data, subset = state=='MT')
summary(regression1)

regression1_controls <- lm(meth~before + age + tv + sex, data = data, subset = state=='MT')
summary(regression1_controls)

```

> From these regressions, we see that the meth usage in Montana before the campaign is 7.6% and went down 3.8% (0.767 - 0.038) after the campaign.  

##1.2 Cross-Sectional Estimator
```{r cross}
mean_MT <- mean(data$state == "MT")
mean_XX <- mean(data$state == "XX")
mean_states <- cbind(mean_MT, mean_XX)

regression2_MT <- lm(meth~before, data = data, subset=state=='MT')
summary(regression2_MT)

regression2_XX <- lm(meth~before, data = data, subset=state=='XX')
summary(regression2_XX)

regression2_MT_control <- lm(meth~before + age + tv + sex, data = data, subset = state=='MT')
summary(regression2_MT_control)

regression2_XX_control <- lm(meth~before + age + tv + sex, data = data, subset = state=='XX')
summary(regression2_XX_control)


```

> In comparison to Montana, across the nation the usage of meth before the campaign was about 6%, and after the campaign there was a 4% decrease in meth usage. 

##1.3 5 Differences-in-Differences
```{r difference}
MT_before <- data %>% filter(data$state=='MT' & data$before==TRUE)
MT_before_mean <- mean(MT_before$meth)

MT_after <- data %>% filter(data$state=='MT' & data$before==FALSE)
MT_after_mean <- mean(MT_after$meth)

XX_before <- data %>% filter(data$state=='XX' & data$before==TRUE)
XX_before_mean <- mean(XX_before$meth)

XX_after <- data %>% filter(data$state=='XX' & data$before==FALSE)
XX_after_mean <- mean(XX_after$meth)

difference <- cbind(MT_before_mean, MT_after_mean, XX_before_mean, XX_after_mean)

regression3 <- lm(meth~state * before, data = data)
summary(regression3)

regression3_controls <- lm(meth~state * before + age + sex + tv, data = data)
summary(regression3_controls)
```

>  From the regression we can see that the other control variables don't seem to play a factor in the meth usage in Montana and nationally. All of the data presented in this Before and After test suggest that the campaign dropped meth usage with both states. 