---
title: "Lab 3 Markdown"
author: "Nicole Bohra"
date: "1/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

##Section 1: 
```{r read, echo=TRUE}
setwd("/Users/nicolebohra/INFO371/Labs")
wvs<- read.delim("wvs.csv.bz2", header = TRUE)
summary(wvs)
```

####Plot: 
```{r plot}
ggplot(wvs) +
  geom_bar(aes(choice, y = ..prop.., fill = country), position = "dodge") +
  scale_x_continuous(breaks = 1:10) +
  labs(y = "frequency", x = "how much freedom of choice do you feel?")
```

####1: Linear model between choice and country
```{r model1}
model <- lm(choice~country, data = wvs)
summary(model)
```

####2: 
> American respondents feel 0.6 units more "freedom" than Chinese 
respondents. In total, Americans are at a Freedom Index of 7.7 in 
responding versus Chinese respondents having about an 
index of 7.1.

####3: 
> The r2 value tells us that 2.5%  of the variance in the data can be explained by the model.

##Section 2: 

####1: Linear model between choice, country and sex
```{r model2}
model_sex <- lm(choice~country + sex, data = wvs)
summary(model_sex)
```

####2: 
> 
a) The level of choice freedom for American men would be about 7.73.
b) The level of choice freedom for American women would be about 7.79.
c) The level of choice freedom for Chinese men would be about 7.13.
d) The level of choice freedom for Chinese women would be about 7.19.


##Section 3: Sex x Country

####1: Linear model with interaction effect
```{r model3}
model_3 <- lm(choice~sex*country, data = wvs)
summary(model_3)
```

####2:
>
a) The level of choice freedom for American men would be about 6.54.
b) The level of choice freedom for American women would be about 7.91.
c) The level of choice freedom for Chinese men would be about 7.26.
d) The level of choice freedom for Chinese women would be about 7.06.

####3: 
> Based on the dataset, it seems that women are more free or have a higher 
 perceived freedom. However, it depends on what country we are looking at.
